---
header-includes:
- \usepackage{amssymb, amsmath, amsthm}
- \usepackage{tabu}
- \newcommand{\E}{\mathbb{E}}
- \newcommand{\var}{{\rm Var}}
- \newcommand{\N}{\mathcal{N}}
output: pdf_document
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{Case 1}           & \\ 
  \textbf{MFE 408: Fixed Income}   & \\ 
  \textbf{Professor Longstaff}         & \\
  \textbf{Group 9}          &\\
  \textbf{Students: Yuhua Deng, Xiahao Wang, Nupur Solanki, Haoxuan Tong}
\end{tabu}

## Arbitrage Case 1

### To find the arbitrage opportunity, we need to find 2 treasury bonds, 1 treasury strip principle bond expiring on the same day.

Step 1:Create a replicated bond of the smaller coupon bond, using the strip coupon bond and the higher coupon bond.

Step 2: Compare the price of the replicated and actual bond, long the cheaper and short the expensive

We will illustrate with one example:

From the slides, we know that current Date: 9th Jan 2015
Bonds with Maturity Date: 11/15/22
Previous coupon payment: 11/15/2014

Treasurey strip: 86.15

Find clean price
Coupon Bond A with rate 1.625: 98.57812
Coupon Bond B with rate 7.625: 143.0625

Find dirty price

```{r}
bondAPrice <- 98.57812
bondARate <- 1.625

bondBPrice <- 143.0625
bondBRate <- 7.625

daysSincePayment <- as.numeric(as.Date('01/09/2015','%m/%d/%Y') - as.Date('11/15/2014','%m/%d/%Y'))
daysInBetweenPayments <- 182

dirty_priceA <- (daysSincePayment/daysInBetweenPayments) * (bondARate/2) + bondAPrice
dirty_priceB <- (daysSincePayment/daysInBetweenPayments) * (bondBRate/2) + bondBPrice
cat("Dirty price for A: ", dirty_priceA,"\n")
cat("Dirty price for B: ", dirty_priceB)
```

To generate a low coupon bond

$$a * 0 + (1 - a) * 7.625 = 1.625$$
```{r}
a <- (7.626 -1.625) / 7.625
cat("a: ",a,"\n")

bondPrice <- a * 86.15 + (1-a) * dirty_priceB
cat("Bond price for the replicated bond is: ",bondPrice,"\n")
cat("This price is lower than the dirty price for bond with coupon rate 1.625, which is ",dirty_priceA)
```
Hence, We long the replicated bond using 0 coupon and 7.625 coupon bond, and short actual 1.625 coupon bond to realize a profit of 0.31




